---
title: "Pneumonia Incidence After CABG Surgery (2015–2021)"
author: "Yixuan He"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: cosmo
    source_code: embed
    self_contained: true
    navbar:
      - { title: "GitHub", icon: "fa-github", href: "https://github.com/yixuan-he425/CABG-pneumonia-dashboard", align: right }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(flexdashboard)
library(plotly)
library(DT)
library(dplyr)
library(tidyr)
library(crosstalk)
```

<style>
/* Force black text in all panels */
.chart-wrapper .chart-shim,
.chart-wrapper .chart-shim * {
  color: #000000 !important;
}
/* Make chart-shim fill its container so flexbox vertical centering works */
.chart-shim {
  height: 100% !important;
  overflow: auto !important;
}
/* Key Findings list */
.chart-shim ul { font-size: 17px !important; line-height: 1.8 !important; }
.chart-shim ul li { font-size: 17px !important; margin-bottom: 6px !important; }
.chart-shim p  { font-size: 17px !important; line-height: 1.8 !important; }

/* Footer (real-world impact) */
.dashboard-footer {
  font-size: 13px;
  color: #000;
  background: #f4f6f8;
  border-top: 2px solid #2E75B6;
  padding: 8px 18px;
  text-align: center;
}
</style>

```{r data-prep}
# ============================================
# DATA FOR VISUALIZATION 1 (incidence.csv)
# ============================================
inc_raw <- read.csv("incidence.csv", fileEncoding = "UTF-8")
names(inc_raw) <- gsub("^\uFEFF", "", names(inc_raw))
names(inc_raw) <- gsub("^X\\.", "", names(inc_raw))

annual <- inc_raw %>%
  group_by(YEAR) %>%
  summarise(
    Patients        = n(),
    Pneumonia_Cases = sum(PNEUMONIA),
    Pneumonia_Rate  = round(Pneumonia_Cases / Patients * 100, 1),
    .groups = "drop"
  ) %>%
  arrange(YEAR) %>%
  mutate(
    Cumulative_Patients = cumsum(Patients),
    Cumulative_Cases    = cumsum(Pneumonia_Cases),
    Cumulative_Rate     = round(Cumulative_Cases / Cumulative_Patients * 100, 1),
    Rate_Change         = round(Pneumonia_Rate - lag(Pneumonia_Rate), 1)
  )

# ============================================
# DATA FOR VISUALIZATION 2 (cohort CSV)
# ============================================
cohort <- read.csv("CABG_pneumonia_cohort_clean_refine.csv")
names(cohort) <- trimws(names(cohort))

# Reverse one-hot: Race
cohort$Race <- case_when(
  cohort$RACE_0 == 1 ~ "White",
  cohort$RACE_1 == 1 ~ "Black or African American",
  cohort$RACE_2 == 1 ~ "Asian",
  cohort$RACE_3 == 1 ~ "Other",
  cohort$RACE_4 == 1 ~ "Unknown/Not Reported",
  TRUE ~ NA_character_
)

# Reverse one-hot: Anesthesia
cohort$Anesthesia <- ifelse(cohort$ANESTHES_General == 1,
                            "General Anesthesia", "Other Anesthesia")

# Reverse one-hot: Diabetes
cohort$Diabetes <- case_when(
  cohort$DIAB_no == 1          ~ "No",
  cohort$DIAB_non.insulin == 1 ~ "Non-insulin",
  cohort$DIAB_insulin == 1     ~ "Insulin",
  TRUE ~ NA_character_
)

# Factor labels
cohort$Gender <- factor(cohort$GENDER, levels = c(0, 1),
                        labels = c("Male", "Female"))

cohort$ASA_Class <- factor(cohort$ASACLAS, levels = 1:5,
  labels = c("ASA 1", "ASA 2", "ASA 3", "ASA 4", "ASA 5"))

cohort$Functional_Status <- factor(cohort$FNSTATUS2, levels = c(0, 1, 2),
  labels = c("Independent", "Partially Dependent", "Totally Dependent"))

# Binary variables
binary_map <- c(
  "SMOKER" = "Smoker", "HXCOPD" = "COPD History", "ASCITES" = "Ascites",
  "HXCHF" = "CHF History", "HYPERMED" = "Hypertension", "DIALYSIS" = "Dialysis",
  "STEROID" = "Steroid Use", "BLEEDIS" = "Bleeding Disorder",
  "DISCANCR" = "Disseminated Cancer", "TRANSFUSION" = "Preop Transfusion"
)

for (orig in names(binary_map)) {
  cohort[[binary_map[[orig]]]] <- factor(cohort[[orig]], levels = c(0, 1),
                                         labels = c("No", "Yes"))
}

cohort$Pneumonia <- factor(cohort$PNEUMONIA, levels = c(0, 1),
                           labels = c("No Pneumonia", "Pneumonia"))

# Color palette
clr_no  <- "#66c2a5"
clr_yes <- "#fc8d62"

# Reusable takeaway HTML
takeaway_eda <- '<div style="display: flex; align-items: flex-start; height: 100%; padding: 5px 15px; overflow-y: auto; overflow-x: hidden;">
<div style="color: #000000; font-size: 13px; line-height: 1.6; text-align: left; border-left: 5px solid #2E75B6; background: #eef2f7; padding: 12px 18px; border-radius: 0 8px 8px 0; width: 100%;">
The exploratory dashboard shows that postoperative pneumonia is a <strong>relatively rare outcome, affecting about 5% of CABG patients</strong>. The study population is mostly older and male, with pneumonia cases appearing across the same broad patient groups seen in the full cohort. Some characteristics — such as <strong>bleeding disorders, preoperative transfusion, and higher BUN values</strong> — appear more common among pneumonia cases, but these patterns are descriptive only and would need further analysis to confirm.
</div>
</div>'
```

Incidence Over Time {data-icon="fa-chart-line"}
=======================================================================

Row {data-height=80}
-----------------------------------------------------------------------

### Total CABG Patients {.value-box}

```{r}
valueBox(format(sum(annual$Patients), big.mark = ","),
         icon = "fa-users", color = "primary")
```

### Overall Pneumonia Rate {.value-box}

```{r}
overall_rate <- round(sum(annual$Pneumonia_Cases) / sum(annual$Patients) * 100, 1)
valueBox(paste0(overall_rate, "%"), icon = "fa-lungs", color = "warning")
```

### Peak Year {.value-box}

```{r}
peak <- annual %>% filter(Pneumonia_Rate == max(Pneumonia_Rate))
valueBox(paste0(peak$YEAR, " (", peak$Pneumonia_Rate, "%)"),
         icon = "fa-arrow-up", color = "danger")
```

### Study Period {.value-box}

```{r}
valueBox("2015 – 2021", icon = "fa-calendar", color = "info")
```

Row {data-height=550}
-----------------------------------------------------------------------

### Annual & Cumulative Pneumonia Incidence

```{r}
hover_annual <- paste0(
  "<b>", annual$YEAR, "</b><br>",
  "Annual Pneumonia Rate: ", annual$Pneumonia_Rate, "%<br>",
  annual$Pneumonia_Cases, " / ", format(annual$Patients, big.mark = ","), " patients",
  ifelse(!is.na(annual$Rate_Change),
         paste0("<br>Change vs prior: ",
                ifelse(annual$Rate_Change > 0, "+", ""), annual$Rate_Change, " pp"),
         "")
)

hover_cum <- paste0(
  "<b>", annual$YEAR, "</b><br>",
  "Cumulative Pneumonia Rate: ", annual$Cumulative_Rate, "%"
)

hover_bar <- paste0(
  "<b>", annual$YEAR, "</b><br>",
  "Cumulative CABG Patients: ", format(annual$Cumulative_Patients, big.mark = ",")
)

covid_row <- annual %>% filter(YEAR == 2020)

plot_ly() %>%
  add_bars(
    data = annual, x = ~YEAR, y = ~Cumulative_Patients,
    name = "Cumulative CABG Patients", marker = list(color = "rgba(200,200,200,0.45)"),
    width = 0.6, yaxis = "y2",
    hovertext = hover_bar, hoverinfo = "text", textposition = "none"
  ) %>%
  add_trace(
    data = annual, x = ~YEAR, y = ~Pneumonia_Rate,
    type = "scatter", mode = "lines+markers",
    name = "Annual Pneumonia Rate (%)",
    line = list(color = "#E74C3C", width = 3),
    marker = list(color = "#E74C3C", size = 9),
    text = hover_annual, hoverinfo = "text"
  ) %>%
  add_trace(
    data = annual, x = ~YEAR, y = ~Cumulative_Rate,
    type = "scatter", mode = "lines+markers",
    name = "Cumulative Pneumonia Rate (%)",
    line = list(color = "#2ECC71", width = 2, dash = "dash"),
    marker = list(color = "#2ECC71", size = 6),
    text = hover_cum, hoverinfo = "text"
  ) %>%
  add_annotations(
    x = annual$YEAR, y = annual$Pneumonia_Rate,
    text = paste0("<b>", annual$Pneumonia_Rate, "%</b>"),
    showarrow = FALSE, yshift = 18,
    font = list(color = "#E74C3C", size = 11)
  ) %>%
  add_annotations(
    x = 2020, y = covid_row$Pneumonia_Rate,
    text = paste0("<b>2020: COVID-era spike</b><br>(+", covid_row$Rate_Change, " pp)"),
    showarrow = TRUE, arrowhead = 2, arrowsize = 1.2,
    arrowcolor = "#C0392B", ax = -80, ay = -50,
    font = list(color = "#C0392B", size = 11),
    bgcolor = "rgba(255,255,255,0.9)",
    bordercolor = "#C0392B", borderwidth = 1, borderpad = 5
  ) %>%
  layout(
    title = list(
    text = "<b>Pneumonia Incidence After <br>CABG Surgery (2015–2021)</b>",
    font = list(size = 18),
    y = 0.95),
    xaxis = list(title = "Year", tickmode = "linear", dtick = 1,
                 range = c(2014.3, 2021.7), tickvals = 2015:2021),
    yaxis = list(title = "Pneumonia Rate (%)", range = c(0, 9), side = "left"),
    yaxis2 = list(title = "Cumulative CABG Patients", overlaying = "y", side = "right",
                  showgrid = FALSE, range = c(0, max(annual$Cumulative_Patients) * 1.15)),
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.18),
    hovermode = "x unified",
    margin = list(l = 70, r = 80, t = 30, b = 90)
  )
```

Row {data-height=300}
-----------------------------------------------------------------------

### Summary Table

```{r}
# NOTE: column names use backticks to allow spaces.
# formatStyle below targets the renamed column "Annual Pneumonia Rate".
tbl <- annual %>%
  select(
    `Year`                        = YEAR,
    `CABG Patients`               = Patients,
    `Pneumonia Cases`             = Pneumonia_Cases,
    `Annual Pneumonia Rate`       = Pneumonia_Rate,
    `Pneumonia Rate Change (pp)`  = Rate_Change,
    `Cum. CABG Patients`          = Cumulative_Patients,
    `Cum. Pneumonia Cases`        = Cumulative_Cases,
    `Cum. Pneumonia Rate`         = Cumulative_Rate
  )

datatable(tbl,
  options = list(pageLength = 10, dom = "t", ordering = TRUE,
                 scrollY = "100%", scrollCollapse = TRUE, paging = FALSE,
                 columnDefs = list(list(className = "dt-center", targets = "_all"))),
  rownames = FALSE, class = "cell-border stripe hover"
) %>%
  formatStyle("Annual Pneumonia Rate",
    background = styleColorBar(c(0, 10), "rgba(231,76,60,0.3)"),
    backgroundSize = "98% 80%", backgroundRepeat = "no-repeat",
    backgroundPosition = "center")
```

### Takeaways
<div style="height: 100%; max-height: 100%; padding: 5px 15px; overflow-y: auto; overflow-x: hidden; box-sizing: border-box;">
<div style="color: #000000; font-size: 13px; line-height: 1.6; text-align: left; border-left: 5px solid #2E75B6; background: #eef2f7; padding: 12px 18px; border-radius: 0 8px 8px 0;">
<strong>The annual pneumonia rate among CABG patients fluctuates between 3.9% and 7.2% from 2015 to 2021</strong>, with no clear long-term upward or downward trend. A noticeable spike occurs in <strong>2020 (7.2%)</strong>, which stands out compared with surrounding years and may reflect disruptions during the COVID-19 era or changes in perioperative risk. Despite these yearly fluctuations, the <strong>overall cumulative incidence remains stable around 5%</strong>, suggesting that the overall pneumonia burden in the cohort was relatively consistent over time.
</div>
</div>


Outcome Distribution {data-icon="fa-pie-chart"}
=======================================================================

Row {data-height=80}
-----------------------------------------------------------------------

### No Pneumonia {.value-box}

```{r}
n_no  <- sum(cohort$Pneumonia == "No Pneumonia")
valueBox(format(n_no, big.mark = ","),
         caption = paste0("No Pneumonia (", round(n_no/nrow(cohort)*100, 1), "%)"),
         icon = "fa-check-circle", color = "success")
```

### Pneumonia {.value-box}

```{r}
n_yes <- sum(cohort$Pneumonia == "Pneumonia")
valueBox(format(n_yes, big.mark = ","),
         caption = paste0("Pneumonia (", round(n_yes/nrow(cohort)*100, 1), "%)"),
         icon = "fa-exclamation-triangle", color = "danger")
```

### Imbalance Ratio {.value-box}

```{r}
valueBox(paste0(round(n_no / n_yes, 1), " : 1"),
         caption = "No PNA : PNA ratio", icon = "fa-balance-scale", color = "info")
```

Row {data-height=550}
-----------------------------------------------------------------------

### Pneumonia Outcome Distribution

```{r}
tbl_pie <- cohort %>%
  count(Pneumonia) %>%
  mutate(pct = round(n / sum(n) * 100, 1))

plot_ly(tbl_pie,
  labels = ~Pneumonia, values = ~n, type = "pie",
  textinfo = "text",
  text = ~paste0(pct, "%\n(", format(n, big.mark = ","), ")"),
  textposition = "inside", textfont = list(size = 14),
  hovertext = ~paste0(Pneumonia, "<br>", format(n, big.mark = ","),
                      " patients (", pct, "%)"),
  hoverinfo = "text",
  marker = list(colors = c(clr_no, clr_yes),
                line = list(color = "#fff", width = 2)),
  pull = c(0, 0.05)
) %>%
  layout(
    title = list(text = "<b>Pneumonia Outcome Distribution</b>", font = list(size = 18)),
    showlegend = TRUE,
    legend = list(font = list(size = 12)),
    margin = list(l = 30, r = 30, t = 50, b = 30)
  )
```

### Key Findings

- **Pneumonia is rare (~5%)** in the CABG cohort, creating a class-imbalanced outcome.
- **Most patients (~95%)** do not develop postoperative pneumonia.
- This **class imbalance matters for modeling**: techniques such as SMOTE, class weighting, or PR-AUC (instead of ROC-AUC) may be needed to build effective predictive models.
- The low event rate also means that even small shifts in risk factor prevalence can noticeably affect the observed pneumonia rate year-to-year.

Row {data-height=150}
-----------------------------------------------------------------------

### Takeaways

```{r, results='asis'}
cat(takeaway_eda)
```


Categorical Variables {data-icon="fa-bar-chart" data-navmenu="Cohort EDA"}
=======================================================================

```{r cat-helper}
# Helper function: create a categorical bar chart
cat_bar <- function(data, var_col, display_name, mode = "stack") {
  df_plot <- data %>%
    filter(!is.na(.data[[var_col]])) %>%
    count(.data[[var_col]], Pneumonia) %>%
    rename(Category = 1)
  
  if (mode == "pct") {
    df_plot <- df_plot %>%
      group_by(Category) %>%
      mutate(pct = round(n / sum(n) * 100, 1)) %>%
      ungroup()
    
    plot_ly(df_plot, x = ~Category, y = ~pct, color = ~Pneumonia,
      colors = c(clr_no, clr_yes), type = "bar",
      hovertext = ~paste0(Category, "<br>", Pneumonia, ": ",
                          pct, "% (n=", format(n, big.mark = ","), ")"),
      hoverinfo = "text"
    ) %>%
      layout(barmode = "stack",
             yaxis = list(title = "Proportion (%)", range = c(0, 105)))
  } else {
    plot_ly(df_plot, x = ~Category, y = ~n, color = ~Pneumonia,
      colors = c(clr_no, clr_yes), type = "bar",
      hovertext = ~paste0(Category, "<br>", Pneumonia, ": ",
                          format(n, big.mark = ",")),
      hoverinfo = "text"
    ) %>%
      layout(barmode = mode, yaxis = list(title = "Count"))
  }
}

cat_styled <- function(p, display_name) {
  p %>% layout(
    title = list(text = paste0("<b>", display_name, " vs Pneumonia</b>"),
                 font = list(size = 18)),
    xaxis = list(title = display_name),
    legend = list(title = list(text = "Outcome")),
    hovermode = "closest",
    margin = list(l = 60, r = 30, t = 50, b = 80)
  )
}
```

Row {.tabset data-height=550}
-----------------------------------------------------------------------

### Gender (Count)

```{r}
cat_bar(cohort, "Gender", "Gender", "stack") %>% cat_styled("Gender")
```

### Gender (%)

```{r}
cat_bar(cohort, "Gender", "Gender", "pct") %>% cat_styled("Gender")
```

### Race (Count)

```{r}
cat_bar(cohort, "Race", "Race", "stack") %>% cat_styled("Race")
```

### Race (%)

```{r}
cat_bar(cohort, "Race", "Race", "pct") %>% cat_styled("Race")
```

### ASA Class (Count)

```{r}
cat_bar(cohort, "ASA_Class", "ASA Class", "stack") %>% cat_styled("ASA Class")
```

### ASA Class (%)

```{r}
cat_bar(cohort, "ASA_Class", "ASA Class", "pct") %>% cat_styled("ASA Class")
```

### Diabetes (Count)

```{r}
cat_bar(cohort, "Diabetes", "Diabetes", "stack") %>% cat_styled("Diabetes")
```

### Diabetes (%)

```{r}
cat_bar(cohort, "Diabetes", "Diabetes", "pct") %>% cat_styled("Diabetes")
```

### Functional Status (Count)

```{r}
cat_bar(cohort, "Functional_Status", "Functional Status", "stack") %>%
  cat_styled("Functional Status")
```

### Functional Status (%)

```{r}
cat_bar(cohort, "Functional_Status", "Functional Status", "pct") %>%
  cat_styled("Functional Status")
```

### Anesthesia (Count)

```{r}
cat_bar(cohort, "Anesthesia", "Anesthesia", "stack") %>% cat_styled("Anesthesia")
```

### Anesthesia (%)

```{r}
cat_bar(cohort, "Anesthesia", "Anesthesia", "pct") %>% cat_styled("Anesthesia")
```

Row {data-height=150}
-----------------------------------------------------------------------

### Takeaways

```{r, results='asis'}
cat(takeaway_eda)
```


Comorbidities {data-icon="fa-heartbeat" data-navmenu="Cohort EDA"}
=======================================================================

Row {.tabset data-height=550}
-----------------------------------------------------------------------

### Smoker (Count)

```{r}
cat_bar(cohort, "Smoker", "Smoker", "stack") %>% cat_styled("Smoker")
```

### Smoker (%)

```{r}
cat_bar(cohort, "Smoker", "Smoker", "pct") %>% cat_styled("Smoker")
```

### COPD (Count)

```{r}
cat_bar(cohort, "COPD History", "COPD History", "stack") %>% cat_styled("COPD History")
```

### COPD (%)

```{r}
cat_bar(cohort, "COPD History", "COPD History", "pct") %>% cat_styled("COPD History")
```

### CHF (Count)

```{r}
cat_bar(cohort, "CHF History", "CHF History", "stack") %>% cat_styled("CHF History")
```

### CHF (%)

```{r}
cat_bar(cohort, "CHF History", "CHF History", "pct") %>% cat_styled("CHF History")
```

### Hypertension (Count)

```{r}
cat_bar(cohort, "Hypertension", "Hypertension", "stack") %>% cat_styled("Hypertension")
```

### Hypertension (%)

```{r}
cat_bar(cohort, "Hypertension", "Hypertension", "pct") %>% cat_styled("Hypertension")
```

### Bleeding Disorder (Count)

```{r}
cat_bar(cohort, "Bleeding Disorder", "Bleeding Disorder", "stack") %>%
  cat_styled("Bleeding Disorder")
```

### Bleeding Disorder (%)

```{r}
cat_bar(cohort, "Bleeding Disorder", "Bleeding Disorder", "pct") %>%
  cat_styled("Bleeding Disorder")
```

### Preop Transfusion (Count)

```{r}
cat_bar(cohort, "Preop Transfusion", "Preop Transfusion", "stack") %>%
  cat_styled("Preop Transfusion")
```

### Preop Transfusion (%)

```{r}
cat_bar(cohort, "Preop Transfusion", "Preop Transfusion", "pct") %>%
  cat_styled("Preop Transfusion")
```

### Dialysis (Count)

```{r}
cat_bar(cohort, "Dialysis", "Dialysis", "stack") %>% cat_styled("Dialysis")
```

### Steroid Use (Count)

```{r}
cat_bar(cohort, "Steroid Use", "Steroid Use", "stack") %>% cat_styled("Steroid Use")
```

### Disseminated Cancer (Count)

```{r}
cat_bar(cohort, "Disseminated Cancer", "Disseminated Cancer", "stack") %>%
  cat_styled("Disseminated Cancer")
```

### Ascites (Count)

```{r}
cat_bar(cohort, "Ascites", "Ascites", "stack") %>% cat_styled("Ascites")
```

Row {data-height=150}
-----------------------------------------------------------------------

### Takeaways

```{r, results='asis'}
cat(takeaway_eda)
```


Continuous Variables {data-icon="fa-chart-area" data-navmenu="Cohort EDA"}
=======================================================================

```{r cont-helper}
# Helper: histogram + boxplot for a continuous variable
cont_plot <- function(data, var_col, display_name, nbins = 30) {
  df <- data %>%
    filter(!is.na(.data[[var_col]])) %>%
    select(value = all_of(var_col), Pneumonia)
  
  hist_no  <- df %>% filter(Pneumonia == "No Pneumonia") %>% pull(value)
  hist_yes <- df %>% filter(Pneumonia == "Pneumonia") %>% pull(value)
  
  p_hist <- plot_ly(alpha = 0.7) %>%
    add_histogram(x = hist_no, name = "No Pneumonia",
                  marker = list(color = clr_no, line = list(color = "black", width = 0.3)),
                  nbinsx = nbins, hoverinfo = "x+y+name") %>%
    add_histogram(x = hist_yes, name = "Pneumonia",
                  marker = list(color = clr_yes, line = list(color = "black", width = 0.3)),
                  nbinsx = nbins, hoverinfo = "x+y+name") %>%
    layout(barmode = "stack",
           xaxis = list(title = display_name),
           yaxis = list(title = "Count"))
  
  p_box <- plot_ly() %>%
    add_boxplot(x = hist_no, name = "No Pneumonia",
                marker = list(color = clr_no), line = list(color = clr_no),
                fillcolor = "rgba(102,194,165,0.4)", hoverinfo = "x+name") %>%
    add_boxplot(x = hist_yes, name = "Pneumonia",
                marker = list(color = clr_yes), line = list(color = clr_yes),
                fillcolor = "rgba(252,141,98,0.4)", hoverinfo = "x+name") %>%
    layout(xaxis = list(showticklabels = FALSE, title = ""),
           yaxis = list(showticklabels = FALSE, title = ""))
  
  subplot(p_box, p_hist, nrows = 2, shareX = TRUE, heights = c(0.25, 0.75)) %>%
    layout(
      title = list(text = paste0("<b>",display_name, " Distribution by Pneumonia Status</b>"),
                   font = list(size = 18)),
      showlegend = TRUE,
      legend = list(title = list(text = "Outcome"), orientation = "v", x = 1.02, y = 0.5),
      margin = list(l = 60, r = 80, t = 50, b = 60),
      hovermode = "closest"
    )
}
```

Row {.tabset data-height=550}
-----------------------------------------------------------------------

### Age (years)

```{r}
cont_plot(cohort, "AGE_num", "Age (years)", 30)
```

### BMI (kg/m²)

```{r}
cont_plot(cohort, "BMI", "BMI (kg/m²)", 30)
```

### WBC (×10⁹/L)

```{r}
cont_plot(cohort, "PRWBC", "WBC (×10⁹/L)", 30)
```

### Hematocrit (%)

```{r}
cont_plot(cohort, "PRHCT", "Hematocrit (%)", 30)
```

### BUN (mg/dL)

```{r}
cont_plot(cohort, "PRBUN", "BUN (mg/dL)", 30)
```

### Creatinine (mg/dL)

```{r}
cont_plot(cohort, "PRCREAT", "Creatinine (mg/dL)", 30)
```

### Serum Sodium (mmol/L)

```{r}
cont_plot(cohort, "PRSODM", "Serum Sodium (mmol/L)", 30)
```

### Albumin (g/dL)

```{r}
cont_plot(cohort, "PRALBUM", "Albumin (g/dL)", 25)
```

### Bilirubin (mg/dL)

```{r}
cont_plot(cohort, "PRBILI", "Bilirubin (mg/dL)", 25)
```

### SGOT/AST (U/L)

```{r}
cont_plot(cohort, "PRSGOT", "SGOT / AST (U/L)", 25)
```

### Platelet Count (×10⁹/L)

```{r}
cont_plot(cohort, "PRPLATE", "Platelet Count (×10⁹/L)", 30)
```

### PTT (seconds)

```{r}
cont_plot(cohort, "PRPTT", "PTT (seconds)", 25)
```

### INR (ratio)

```{r}
cont_plot(cohort, "PRINR", "INR (ratio)", 25)
```

Row {data-height=150}
-----------------------------------------------------------------------

### Takeaways

```{r, results='asis'}
cat(takeaway_eda)
```


Dataset {data-icon="fa-database"}
=======================================================================

Row {data-height=850}
-----------------------------------------------------------------------

### About the Data

<div style="padding: 25px 35px; font-size: 15px; line-height: 1.8; color: #000000;">

<h3 style="color: #2E75B6; border-bottom: 2px solid #2E75B6; padding-bottom: 6px;">
Data Source &amp; Description
</h3>

<p>
This dashboard uses data from the <strong>American College of Surgeons National Surgical Quality Improvement Program (ACS-NSQIP)</strong> Participant Use Data Files for 2015–2021. General information about ACS-NSQIP is available from the American College of Surgeons website:
<a href="https://www.facs.org/quality-programs/data-and-registries/acs-nsqip/" target="_blank">
https://www.facs.org/quality-programs/data-and-registries/acs-nsqip/
</a>.
The analytic cohort includes <strong>13,650 adult patients who underwent isolated coronary artery bypass grafting (CABG)</strong>, identified by relevant CPT codes, with a postoperative pneumonia event rate of about 5.3%.
Data were prospectively collected by trained Surgical Clinical Reviewers at participating U.S. hospitals through standardized chart abstraction of preoperative and 30-day postoperative variables.
The study population consists of adult CABG patients across U.S. ACS-NSQIP-participating hospitals between January 1, 2015 and December 31, 2021.
</p>

<h3 style="color: #C0392B; border-bottom: 2px solid #C0392B; padding-bottom: 6px; margin-top: 20px;">
Data Availability &amp; Confidentiality
</h3>

<div style="background: #fdecea; border-left: 5px solid #C0392B; padding: 14px 20px; border-radius: 0 8px 8px 0;">
<p style="margin: 0;">
<strong>The ACS-NSQIP dataset used in this dashboard is confidential and cannot be shared publicly.</strong>
Although ACS-NSQIP data are de-identified, access requires a formal institutional Data Use Agreement (DUA) with the American College of Surgeons.
<strong>Raw patient-level data are not posted in the GitHub repository and are not redistributable.</strong>
Only aggregated, non-identifying summary statistics are displayed in this dashboard.
The rendered static HTML dashboard can be viewed without running the project code, but reproducing the dashboard or the complete machine-learning workflow requires access to the original ACS-NSQIP dataset. Researchers wishing to reproduce the dashboard must obtain the dataset directly through their own institution’s NSQIP DUA.
</p>
</div>

<h3 style="color: #2E75B6; border-bottom: 2px solid #2E75B6; padding-bottom: 6px; margin-top: 20px;">
Variables Shown in This Dashboard
</h3>

<p><strong>Outcome</strong></p>
<ul>
<li>30-day postoperative pneumonia (binary)</li>
</ul>

<p><strong>Demographics</strong></p>
<ul>
<li>Age (years)</li>
<li>Gender</li>
<li>Race / Ethnicity</li>
<li>Body Mass Index (BMI, kg/m²)</li>
</ul>

<p><strong>Surgical &amp; Anesthesia Context</strong></p>
<ul>
<li>Anesthesia Technique</li>
<li>American Society of Anesthesiologists Physical Status (ASA Class)</li>
<li>Functional Health Status</li>
<li>Preoperative Red Blood Cell (RBC) Transfusion</li>
</ul>

<p><strong>Comorbidities</strong></p>
<ul>
<li>Diabetes Mellitus</li>
<li>Current Smoker</li>
<li>History of Severe Chronic Obstructive Pulmonary Disease (COPD)</li>
<li>History of Congestive Heart Failure (CHF)</li>
<li>Hypertension Requiring Medication</li>
<li>Currently on Dialysis</li>
<li>Steroid / Immunosuppressant Use</li>
<li>Bleeding Disorders</li>
<li>Disseminated Cancer</li>
<li>Ascites</li>
</ul>

<p><strong>Preoperative Laboratory Values</strong></p>
<ul>
<li>Serum Sodium (mEq/L)</li>
<li>Blood Urea Nitrogen (BUN, mg/dL)</li>
<li>Serum Creatinine (mg/dL)</li>
<li>Albumin (g/dL)</li>
<li>Total Bilirubin (mg/dL)</li>
<li>Aspartate Aminotransferase / Serum Glutamic-Oxaloacetic Transaminase (AST/SGOT, U/L)</li>
<li>White Blood Cell Count (WBC, K/mm³)</li>
<li>Hematocrit (HCT, %)</li>
<li>Platelet Count (K/cumm)</li>
<li>Partial Thromboplastin Time (PTT, seconds)</li>
<li>International Normalized Ratio (INR, unitless)</li>
</ul>

<h3 style="color: #2E75B6; border-bottom: 2px solid #2E75B6; padding-bottom: 6px; margin-top: 20px;">
Notes &amp; Limitations
</h3>

<ul>
<li>All visualizations describe associations only; <strong>no causal inference</strong> should be drawn from this dashboard.</li>
<li>Pneumonia is a <strong>rare event (~5%)</strong>, so apparent group differences should be interpreted with caution.</li>
<li>Year-to-year variation may reflect changes in case mix, hospital participation, or external factors such as the COVID-19 pandemic in 2020.</li>
</ul>

</div>


About {data-icon="fa-info-circle"}
=======================================================================

Row {data-height=850}
-----------------------------------------------------------------------

### Project Overview

<div style="padding: 25px 35px; font-size: 15px; line-height: 1.8; color: #000000;">

<h3 style="color: #2E75B6; border-bottom: 2px solid #2E75B6; padding-bottom: 6px;">
Why This Project Matters
</h3>

<p>
Although postoperative pneumonia after CABG is a leading driver of prolonged ventilation, ICU stay, and 30-day mortality, its overall patterns, temporal trends, and patient risk-factor distributions remain insufficiently described in large contemporary U.S. surgical datasets.
By visualizing temporal trends and patient-level distributions across 13,650 CABG cases in ACS-NSQIP (2015–2021), this dashboard supports clinicians, quality-improvement teams, and researchers in identifying high-risk subgroups and informing predictive modeling efforts.
</p>

<h3 style="color: #2E75B6; border-bottom: 2px solid #2E75B6; padding-bottom: 6px; margin-top: 20px;">
Source Code &amp; Repository
</h3>

<p>
The complete dashboard source code, rendered HTML file, and project README are available in the GitHub repository:
<a href="https://github.com/yixuan-he425/CABG-pneumonia-dashboard" target="_blank">
https://github.com/yixuan-he425/CABG-pneumonia-dashboard
</a>.
</p>

<h3 style="color: #2E75B6; border-bottom: 2px solid #2E75B6; padding-bottom: 6px; margin-top: 20px;">
Author
</h3>

<p>
Yixuan He, MSPH/MPH Candidate, Department of Biostatistics &amp; Bioinformatics, Rollins School of Public Health, Emory University.
</p>

</div>
